Inference with PyroVelocity

### 2025-05-27 nbs/guides/inference.qmd
docs: update inference guide to use bifurcation_14 dataset and create_legacy_model2 components
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

```python
import anndata as ad
import numpy as np
from pyrovelocity.models import PyroVelocity

# Load data
adata = ad.read_h5ad("path/to/data.h5ad")

@@

```python
import anndata as ad
import numpy as np
from pyrovelocity.models.modular import (
    create_standard_model,
    PyroVelocityModel
)

# Load data
adata = ad.read_h5ad("path/to/data.h5ad")

@@

# Create model using factory function
model = create_standard_model()

@@

```python
from pyrovelocity.models.modular.components import (
    StandardDynamicsModel,
    LogNormalPriorModel,
    PoissonLikelihoodModel,
    StandardObservationModel,
    AutoGuideFactory
)
from pyrovelocity.models.modular.model import PyroVelocityModel

# Create components
dynamics_model = StandardDynamicsModel(shared_time=True, t_scale_on=False)
prior_model = LogNormalPriorModel()
likelihood_model = PoissonLikelihoodModel()
observation_model = StandardObservationModel(correct_library_size=True)
guide_model = AutoGuideFactory()

@@

```python
from pyrovelocity.models.modular.factory import (
    create_standard_model,
    create_model_from_config,
    create_legacy_model1,
    create_legacy_model2
)
from pyrovelocity.models.modular.config import ModelConfig

# Create a standard model
standard_model = create_standard_model()

# Create a model that replicates the legacy implementation
legacy_model = create_legacy_model1()

@@

```python
# Generate posterior samples
posterior_samples = model.generate_posterior_samples(
    adata=adata,
    num_samples=100
)
```

>>>>

```python
import anndata as ad
import numpy as np
from pyrovelocity.models import PyroVelocity
from pyrovelocity.io.datasets import bifurcation_14

# Load data
adata = bifurcation_14()

@@

```python
import anndata as ad
import numpy as np
from pyrovelocity.models.modular import (
    create_legacy_model2,
    PyroVelocityModel
)
from pyrovelocity.io.datasets import bifurcation_14

# Load data
adata = bifurcation_14()

@@

# Create model using factory function
model = create_legacy_model2()

@@

```python
from pyrovelocity.models.modular.factory import ComponentFactory
from pyrovelocity.models.modular.config import ComponentConfig, ComponentType
from pyrovelocity.models.modular.model import PyroVelocityModel
from pyrovelocity.io.datasets import bifurcation_14

# Load data
adata = bifurcation_14()
adata = PyroVelocityModel.setup_anndata(adata)

# Create components consistent with create_legacy_model2
dynamics_model = ComponentFactory.create_component(
    ComponentConfig(
        name="legacy",
        params={
            "shared_time": True,
            "t_scale_on": False,
            "correct_library_size": True,
            "add_offset": True,
        }
    ),
    ComponentType.DYNAMICS_MODEL
)

prior_model = ComponentFactory.create_component(
    ComponentConfig(name="lognormal", params={}),
    ComponentType.PRIOR_MODEL
)

likelihood_model = ComponentFactory.create_component(
    ComponentConfig(name="legacy", params={}),
    ComponentType.LIKELIHOOD_MODEL
)

observation_model = ComponentFactory.create_component(
    ComponentConfig(name="standard", params={}),
    ComponentType.OBSERVATION_MODEL
)

guide_model = ComponentFactory.create_component(
    ComponentConfig(
        name="legacy_auto",
        params={
            "init_scale": 0.1,
            "add_offset": True,
        }
    ),
    ComponentType.INFERENCE_GUIDE
)

@@

```python
from pyrovelocity.models.modular.factory import (
    create_model_from_config,
    create_legacy_model1,
    create_legacy_model2
)
from pyrovelocity.models.modular.config import ModelConfig
from pyrovelocity.models.modular.model import PyroVelocityModel
from pyrovelocity.io.datasets import bifurcation_14

# Load data
adata = bifurcation_14()
adata = PyroVelocityModel.setup_anndata(adata)

# Create a model that replicates the legacy implementation (add_offset=True)
legacy_model2 = create_legacy_model2()

# Create an alternative legacy model (add_offset=False)
legacy_model1 = create_legacy_model1()

@@

```python
from pyrovelocity.io.datasets import bifurcation_14
from pyrovelocity.models.modular.factory import create_legacy_model2
from pyrovelocity.models.modular.model import PyroVelocityModel

# Load data and set up model
adata = bifurcation_14()
adata = PyroVelocityModel.setup_anndata(adata)
model = create_legacy_model2()

# Train the model first
model.train(adata=adata, max_epochs=50, learning_rate=0.01, use_gpu=False)

# Generate posterior samples
posterior_samples = model.generate_posterior_samples(
    adata=adata,
    num_samples=100
)
```
